---
title: "The association of the 24-hour activity cycle profiles with cognitive function in older adults with Mild Cognitive Impairment: A cross-sectional study"
author: "Guilherme Moraes Balbim"
date: "08/25/2023"
output: 
  pdf_document: 
    toc: yes
    number_sections: yes
    toc_depth: 5
  word_document: 
    toc: yes
    toc_depth: 5
geometry: "left = 1cm, right = 1cm, top = 1cm, bottom = 2.5cm"
header-includes:
- \usepackage{caption}
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
editor_options: 
  chunk_output_type: console
---

\newpage

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  include = FALSE,
  fig.height = 7, fig.width = 9, fig.align = "center")
```

\small
# Loading packages
```{r message=FALSE}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse,dplyr,tidyr,lubridate,knitr,openxlsx,readxl,tidyLPA,mclust,tableone,tinytex,compositions,ggmoldes,mgcv,skimr,GGally,rstatix,broom,psych,performance,janitor,ggplot2, PerformanceAnalytics, emmeans)

```

# Loading data for initial cleaning and descriptives
```{r message=FALSE}

data_of_interest <- read_csv("../03.spreadsheets/all_data_merged_averages_latentcog_20230131.csv", 
                         show_col_types = FALSE,
                         name_repair = make_clean_names,
                         col_select = c(1:11, 13, 16:19, 23:27, 49))

```

# Descriptives
## Sleep, PA, SB 
```{r}
# Create a study variable
data_of_interest <- data_of_interest %>% 
  mutate(rct = str_extract(id, "[:alpha:]"))

# Exclude participants with <5 days of data, < 4h, and > 12h of sleep = 240 participants

data_of_interest <- data_of_interest %>% 
  filter(total_days >= 5) %>% 
  filter(avg_sleep_min > 240 | avg_sleep_min < 720)

skim(data_of_interest)

# Differences by RCT
kable(print(printToggle = FALSE, CreateTableOne(data = data_of_interest, 
                                                c("avg_sleep_min",
                                                  "avg_sedentary_min",
                                                  "avg_light_pa_min",
                                                  "avg_mvpa_min",
                                                  "avg_sleep_pct",
                                                  "avg_sedentary_pct",
                                                  "avg_light_pa_pct",
                                                  "avg_mvpa_pct"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE))) # NS

## Histograms
data_of_interest %>%
  dplyr::select("avg_sleep_min","avg_sedentary_min","avg_light_pa_min","avg_mvpa_min","avg_sleep_pct","avg_sedentary_pct","avg_light_pa_pct","avg_mvpa_pct") %>%
  pivot_longer(cols = 1:8, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  facet_wrap("measures", scale = "free") +
  geom_histogram(colour = "black", binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3))) + # Proportional bins
  theme_minimal() +
  theme(legend.position = "none")


## Bloxplot
data_of_interest %>%
  dplyr::select("avg_sleep_min","avg_sedentary_min","avg_light_pa_min","avg_mvpa_min","avg_sleep_pct","avg_sedentary_pct","avg_light_pa_pct","avg_mvpa_pct") %>%
  pivot_longer(cols = 1:8, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  geom_boxplot(colour = "black") + # Proportional bins
  ylim(-2,2) +
  facet_wrap("measures", scale = "free", ncol = 2) +
  coord_flip() +
  theme_minimal() +
  theme(legend.position = "none") # Some potential outliers
```

## Demographics
```{r}
# Continuous data
## Differences by RCT
kable(print(printToggle = FALSE, CreateTableOne(data = data_of_interest, 
                                                c("age",
                                                  "moca",
                                                  "mmse",
                                                  "height",
                                                  "weight",
                                                  "bmi"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE)))

# Categorical data
## Differences by RCT
kable(print(printToggle = FALSE, CreateTableOne(data = data_of_interest, 
                                                c("sex", "education", "bmi_cat"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE)))

## Normality tests
kable(digits = 3, caption = "Normality test", data_of_interest %>% 
        shapiro_test(avg_sleep_min,
                     avg_sedentary_min,
                     avg_light_pa_min,
                     avg_mvpa_min,
                     avg_sleep_pct,
                     avg_sedentary_pct,
                     avg_light_pa_pct,
                     avg_mvpa_pct)) # MVPA non-normal
```

## Outliers
```{r}
## Checking outliers
avg_sleep_pct_outliers <- check_outliers(data_of_interest$avg_sleep_pct)
avg_sleep_pct_outliers # 3 outliers detected: cases 54, 127, 233.

avg_sedentary_pct_outliers <- check_outliers(data_of_interest$avg_sedentary_pct)
avg_sedentary_pct_outliers # 1 outlier detected: case 104.

avg_light_pa_pct_outliers <- check_outliers(data_of_interest$avg_light_pa_pct)
avg_light_pa_pct_outliers # 1 outlier detected: case 127.

avg_mvpa_pct_outliers <- check_outliers(data_of_interest$avg_mvpa_pct)
avg_mvpa_pct_outliers # 14 outliers detected: cases 13, 24, 44, 66, 88, 100, 104, 120, 150, 200, 219, 230, 231, 232.

# Winsorizing outliers
timeusebehav_wins <- data_of_interest

for(var in c("avg_sleep_pct","avg_sedentary_pct","avg_light_pa_pct","avg_mvpa_pct")){
  Q <- quantile(timeusebehav_wins[[var]], probs=c(.05, .95), na.rm = TRUE)
  Q1 <- quantile(timeusebehav_wins[[var]], probs=c(.25), na.rm = TRUE)
  Q3 <- quantile(timeusebehav_wins[[var]], probs=c(.75), na.rm = TRUE)
  iqr <- IQR(timeusebehav_wins[[var]], na.rm = TRUE)
  up <-  Q3+1.5*iqr # Upper Range  
  low <- Q1-1.5*iqr # Lower Range
  new <- timeusebehav_wins %>% 
    mutate(wins = ifelse(timeusebehav_wins[[var]] <= (Q1 - 1.5*iqr), Q[1], 
                         ifelse(timeusebehav_wins[[var]] >= (Q3 + 1.5*iqr), Q[2], timeusebehav_wins[[var]]))) %>% 
    select(wins) %>% 
    rename_with(~paste0(var, "_",.))
  timeusebehav_wins <- cbind(timeusebehav_wins, new)
}

## Checking distribution with winsorized data
### Histograms
timeusebehav_wins %>%
  select("avg_sleep_pct_wins","avg_sedentary_pct_wins","avg_light_pa_pct_wins","avg_mvpa_pct_wins") %>%
  pivot_longer(cols = 1:4, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  facet_wrap("measures", scale = "free") +
  geom_histogram(colour = "black", binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3))) + # Proportional bins
  theme_minimal() +
  theme(legend.position = "none")

## Checking normality of different types of transformations for non-normal
timeusebehav_wins %>%
  select(c("avg_sleep_pct_wins","avg_sedentary_pct_wins","avg_light_pa_pct_wins","avg_mvpa_pct_wins")) %>%
  pivot_longer(cols = 1:4, names_to = "measures", values_to = "data") %>%
  group_by(measures) %>% 
  shapiro_test(data) %>% 
  kable(digits = 3) # MVPA still is non-normal

## 24HAC behaviours by study
kable(print(printToggle = FALSE, CreateTableOne(data = timeusebehav_wins, 
                                                c("avg_sleep_pct_wins",
                                                  "avg_sedentary_pct_wins",
                                                  "avg_light_pa_pct_wins",
                                                  "avg_mvpa_pct_wins"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE))) # NS

data_of_interest <- timeusebehav_wins

```

## ADAS-Cog
```{r}

kable(print(printToggle = FALSE, CreateTableOne(data = data_of_interest, 
                                                c("adas_cog_plus"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE)))

## Histograms
data_of_interest %>%
  select(c("adas_cog_plus")) %>%
  pivot_longer(cols = 1:1, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  facet_wrap("measures", scale = "free") +
  geom_histogram(colour = "black", binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3))) + # Proportional bins
  theme_minimal() +
  theme(legend.position = "none")


## Bloxplot
data_of_interest %>%
  select(c("adas_cog_plus")) %>%
  pivot_longer(cols = 1, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  geom_boxplot(colour = "black") + # Proportional bins
  ylim(-2,2) +
  facet_wrap("measures", scale = "free", ncol = 2) +
  coord_flip() +
  theme_minimal() +
  theme(legend.position = "none") # Some potential outliers

# Normality tests
kable(digits = 3, caption = "Normality test", data_of_interest %>% 
        shapiro_test(adas_cog_plus)) # Non-normal

# Checking outliers 
adas_cog_plus_outliers <- check_outliers(data_of_interest$adas_cog_plus, method = "iqr")
adas_cog_plus_outliers # 4 outliers detected (cases 99, 129, 152, 222).

# Winsorizing outliers
data_of_interest_wins <- data_of_interest

for(var in c("adas_cog_plus")){
  Q <- quantile(data_of_interest_wins[[var]], probs=c(.05, .95), na.rm = TRUE)
  Q1 <- quantile(data_of_interest_wins[[var]], probs=c(.25), na.rm = TRUE)
  Q3 <- quantile(data_of_interest_wins[[var]], probs=c(.75), na.rm = TRUE)
  iqr <- IQR(data_of_interest_wins[[var]], na.rm = TRUE)
  up <-  Q3+1.5*iqr # Upper Range  
  low <- Q1-1.5*iqr # Lower Range
  new <- data_of_interest_wins %>% 
    mutate(wins = ifelse(data_of_interest_wins[[var]] <= (Q1 - 1.5*iqr), Q[1], 
                         ifelse(data_of_interest_wins[[var]] >= (Q3 + 1.5*iqr), Q[2], data_of_interest_wins[[var]]))) %>% 
    select(wins) %>% 
    rename_with(~paste0(var, "_",.))
  data_of_interest_wins <- cbind(data_of_interest_wins, new)
}

## Checking distribution with winsorized data
### Histograms
data_of_interest_wins %>%
  select("adas_cog_plus_wins") %>%
  pivot_longer(cols = 1:1, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  facet_wrap("measures", scale = "free") +
  geom_histogram(colour = "black", binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3))) + # Proportional bins
  theme_minimal() +
  theme(legend.position = "none")

## Checking normality of different types of transformations for non-normal
data_of_interest_wins %>%
  select(c("adas_cog_plus_wins")) %>%
  pivot_longer(cols = 1:1, names_to = "measures", values_to = "data") %>%
  mutate(data_log = log10(data),
         data_sqrt = sqrt(data)) %>% 
  group_by(measures) %>% 
  shapiro_test(data_log,
               data_sqrt,
               data) %>% 
  kable(digits = 3) # ADAS-Cog without transformation looks better

## Checking ADAS-Cog by study
kable(print(printToggle = FALSE, CreateTableOne(data = data_of_interest_wins, 
                                                c("adas_cog_plus_wins"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE)))

# Saving the dataset 
# write.csv(data_of_interest_wins, "../03.spreadsheets/all_data_merged_clean_wins_20230609.csv")

```

# Loading data and preparing dataset for compositions
```{r message=FALSE}

final_data <- read_csv("../03.spreadsheets/all_data_merged_clean_wins_20230609.csv", 
                       show_col_types = FALSE,
                       name_repair = make_clean_names)
skim(final_data)

```

## Check time-use data adds up 100%
```{r message=FALSE}

rowSums(final_data[, 
                  c("avg_mvpa_pct_wins", "avg_sedentary_pct_wins", "avg_light_pa_pct_wins", "avg_sleep_pct_wins")
]) # No, will need the closure from compositions
  
```

## Closure function (compositions package)
```{r message=FALSE}

# Create dataframes containing the outcome, predictors and covariates
cols_outcome <- c("adas_cog_plus_wins")
cols_pred_pct <- c("avg_mvpa_pct_wins", "avg_sedentary_pct_wins", "avg_light_pa_pct_wins", "avg_sleep_pct_wins")
cols_id_covars <- c("id","sex","age", "bmi", "bmi_cat", "moca", "mmse", "education", "rct")

# Combine all dataframes
cols_pairwise <- c(cols_outcome, cols_pred_pct, cols_id_covars)

# Create new dataframe containing all rows from final_data but only cols_pairwise columns
pairwise_data <- final_data[, cols_pairwise] %>% 
  relocate(id)

# Ensuring that time-use data have a total of 100%
hac_behav_pct_wins <- pairwise_data %>% 
  select("avg_mvpa_pct_wins", "avg_sedentary_pct_wins", "avg_light_pa_pct_wins", "avg_sleep_pct_wins")

hac_behav_pct_clo <- data.frame(clo(hac_behav_pct_wins, total = 100))

rowSums(hac_behav_pct_clo)  # All participants with 100%

# Appending 24hac closure behaviours to final data
hac_behav_pct_clo_id <- bind_cols(final_data$id, hac_behav_pct_clo)
hac_behav_pct_clo_id <- hac_behav_pct_clo_id %>% 
  rename("id" = "...1")

final_data_1 <- left_join(final_data, hac_behav_pct_clo_id, by = "id")
final_data_1 <- final_data_1 %>% 
  rename_with(~gsub("wins.y", "wins_clos", .), contains("wins.y")) %>% 
  rename_with(~gsub(".x", "_noclos", .), contains("wins.x"))

final_data <- final_data_1

```

## Arranging data
```{r message=FALSE}

cols_outcome2 <- c("adas_cog_plus_wins")
cols_pred_pct2 <- c("avg_mvpa_pct_wins_clos", "avg_sedentary_pct_wins_clos", "avg_light_pa_pct_wins_clos",
                    "avg_sleep_pct_wins_clos")
cols_pred_min2 <- c("avg_mvpa_min_wins_clos", "avg_sedentary_min_wins_clos", "avg_light_pa_min_wins_clos",
                    "avg_sleep_min_wins_clos")
cols_pairwise2 <- c(cols_outcome2, cols_pred_pct2, cols_pred_min2)

# Create new dataframe for correlation analysis
correlation_data <- final_data[, cols_pairwise2]

```

## Pearson correlations
```{r message=FALSE}

cor.mat <- correlation_data %>% 
  cor_mat()
cor.mat

cor.pmat <- correlation_data %>% 
  cor_pmat()
cor.pmat # Correlation between time-use and ADAS-Cog were significant for MVPA and SB, but not for LPA and sleep

chart.Correlation(correlation_data, histogram = TRUE, method = "pearson")

```

# Compositional data analysis
Setting up the data to create the ilrs so that the first ilrs ilr1 = MVPA:all remaining, ilr2 = SB:LPA+sleep, ilr3 = LPA:sleep. All three ilrs will be entered in to LPA analysis to capture the entire 24-hour time-use composition.
```{r,include=TRUE}

## Create a matrix
data_comp_mean_min <- matrix(c(final_data$avg_mvpa_min_wins_clos, 
                               final_data$avg_sedentary_min_wins_clos, 
                               final_data$avg_light_pa_min_wins_clos, 
                               final_data$avg_sleep_min_wins_clos), ncol = 4)

data_min_clos <- final_data %>% 
  select("avg_mvpa_min_wins_clos", "avg_sedentary_min_wins_clos", 
         "avg_light_pa_min_wins_clos", "avg_sleep_min_wins_clos")

rowSums(data_min_clos)

data_comp_mean_pct <- matrix(c(final_data$avg_mvpa_pct_wins_clos,
                               final_data$avg_sedentary_pct_wins_clos,
                               final_data$avg_light_pa_pct_wins_clos,
                               final_data$avg_sleep_pct_wins_clos), ncol = 4)

data_pct_clos <- final_data %>% 
  select("avg_mvpa_pct_wins_clos", "avg_sedentary_pct_wins_clos", 
         "avg_light_pa_pct_wins_clos", "avg_sleep_pct_wins_clos")

rowSums(data_pct_clos)

## ILR coordinates - sequential binary partition process
sbp4_3 = matrix(c( 1, -1, -1,-1, 
                 0, 1, -1, -1,
                 0, 0, 1, -1),
              ncol=4, byrow=TRUE)


psi4_3 = gsi.buildilrBase(t(sbp4_3))

## With 3 ilrs. Obs: "avg_mvpa_min_wins_clos", "avg_sedentary_min_wins_clos", "avg_light_pa_min_wins_clos", "avg_sleep_min_wins_clos"

ilrs3 = ilr(acomp(final_data[, (cols_pred_pct2)]), V=psi4_3) # compute ilrs of all rows but only columns in cols_pred
head(ilrs3)

# rename ilrs to ilr1, ilr2 and ilr3
colnames(ilrs3) <- paste0("ilr", 1:3)
head(ilrs3)
nrow(ilrs3)
nrow(final_data) # Number of rows for ilrs and final_data_comp matches

# Get ilrs means
ilr_mean_min <- colMeans(ilrs3)
ilr_mean_min

```

# Descriptives of 24HAC behaviours (compositional means of whole population)
```{r,include=TRUE}
tbl1 <- CreateTableOne(data = final_data, vars = c('sex',
                                                   'avg_sleep_min_wins_clos',
                                                   'avg_sleep_pct_wins_clos',
                                                   'avg_sedentary_min_wins_clos',
                                                   'avg_sedentary_pct_wins_clos',
                                                   'avg_light_pa_min_wins_clos',
                                                   'avg_light_pa_pct_wins_clos',
                                                   'avg_mvpa_min_wins_clos',
                                                   'avg_mvpa_pct_wins_clos')) %>% 
  print(., printToggle = FALSE)

tbl1  
  
rownames(tbl1)[2:10] <- c('Sex = Male (%)',
                          'Sleep (min/day), mean (SD)',
                         'Sleep (%/day), mean (SD)',
                         'Sedentary behaviour (min/day), mean (SD)',
                         'Sedentary behaviour (%/day), mean (SD)',
                         'Light PA (min/day), mean (SD)',
                         'Light PA (%/day), mean (SD)',
                         'Moderate-to-vigorous PA (min/day), mean (SD)',
                         'Moderate-to-vigorous PA (%/day), mean (SD)')

kable(tbl1, align = 'c', vline = '',
      booktabs = TRUE,
      caption = "Descriptive statistics of the sample",
      linesep = '')

```

## Create new dataframe containing the data to be used in LPA
```{r message=FALSE}

ilr3_data <- cbind(final_data, ilrs3) %>% 
  select(-c(1,23)) %>% 
  rename_all( ~ str_replace(., "_wins", "")) %>%
  rename_all( ~ str_replace(., "avg_", "")) %>% 
  rename("ilr_mvpa_others" = "ilr1",
         "ilr_sb_lpa_sleep" = "ilr2",
         "ilr_lpa_sleep" = "ilr3")
            
# Double-check there are no NAs in the data
table(is.na(ilr3_data)) # No NAs
skim(ilr3_data)

# Save dataset
# write.csv(ilr3_data, "../03.spreadsheets/all_data_3ilrs_202300508.csv")

```

# Latent Profile Analysis with 3 ilrs
## Checking normality and dealing with outliers
```{r message=FALSE}

## Histograms
ilr3_data %>%
  select(c("ilr_mvpa_others", "ilr_sb_lpa_sleep", "ilr_lpa_sleep")) %>%
  pivot_longer(cols = 1:3, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  facet_wrap("measures", scale = "free") +
  geom_histogram(colour = "black", binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3))) + # Proportional bins
  theme_minimal() +
  theme(legend.position = "none")

# Checking outliers 
ilr1_outliers <- check_outliers(ilr3_data$ilr_mvpa_others, method = "iqr")
ilr1_outliers # 3 outliers detected: cases 39, 127, 205.

ilr2_outliers <- check_outliers(ilr3_data$ilr_sb_lpa_sleep, method = "iqr")
ilr2_outliers # No outliers detected

ilr3_outliers <- check_outliers(ilr3_data$ilr_lpa_sleep, method = "iqr")
ilr3_outliers # 3 outliers detected: cases 106, 205, 209.

# Winsorizing outliers
ilrs_wins <- ilr3_data

for(var in c("ilr_mvpa_others", "ilr_sb_lpa_sleep", "ilr_lpa_sleep")){
  Q <- quantile(ilrs_wins[[var]], probs=c(.05, .95), na.rm = TRUE)
  Q1 <- quantile(ilrs_wins[[var]], probs=c(.25), na.rm = TRUE)
  Q3 <- quantile(ilrs_wins[[var]], probs=c(.75), na.rm = TRUE)
  iqr <- IQR(ilrs_wins[[var]], na.rm = TRUE)
  up <-  Q3+1.5*iqr # Upper Range  
  low <- Q1-1.5*iqr # Lower Range
  new <- ilrs_wins %>% 
    mutate(wins = ifelse(ilrs_wins[[var]] <= (Q1 - 1.5*iqr), Q[1], 
                         ifelse(ilrs_wins[[var]] >= (Q3 + 1.5*iqr), Q[2], ilrs_wins[[var]]))) %>% 
    select(wins) %>% 
    rename_with(~paste0(var, "_",.))
  ilrs_wins <- cbind(ilrs_wins, new)
}

## Checking distribution with winsorized data
### Histograms
ilrs_wins %>%
  select("ilr_mvpa_others_wins", "ilr_sb_lpa_sleep_wins", "ilr_lpa_sleep_wins") %>%
  pivot_longer(cols = 1:3, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  facet_wrap("measures", scale = "free") +
  geom_histogram(colour = "black", binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3))) + # Proportional bins
  theme_minimal() +
  theme(legend.position = "none") # All good now

# Checking outliers 
ilr1_outliers <- check_outliers(ilrs_wins$ilr_mvpa_others_wins, method = "iqr")
ilr1_outliers # No outliers

ilr2_outliers <- check_outliers(ilrs_wins$ilr_sb_lpa_sleep_wins, method = "iqr")
ilr2_outliers # No outliers

ilr3_outliers <- check_outliers(ilrs_wins$ilr_lpa_sleep_wins, method = "iqr")
ilr3_outliers # No outliers

# Dataset with only wins ilrs
ilrs_wins <- ilrs_wins %>% 
  select(-c(40:42))

ilrs_wins <-  ilrs_wins %>% 
  rename_all( ~ str_replace(., "_wins", ""))
skim(ilrs_wins)

```

## Exploring LPAs
```{r }

# Comparing solutions from 2 to 6 classes

suppressMessages(mod <- estimate_profiles(df = ilrs_wins[,32:34],n_profiles = 2:6, models = 2))
tbl_lpa <- sapply(mod,function(x) {x$fit}) %>% t(.)
tbl_lpa[,3:17] <- round(tbl_lpa[,3:17],2)
tbl_lpa_to_present <- tbl_lpa[,c(2,3,4,6,7,10,11,15,18,12)]
rownames(tbl_lpa_to_present) <- c('2-class model','3-class model','4-class model','5-class model','6-class model')
tbl_lpa_to_present

```
By fit-model values solution with 4 profiles look the best one, followed by 5 and 3.

### Checking the model (ilr) with 6 classes
```{r }

ilr_model2_6classes <- ilrs_wins %>% 
  dplyr::select(ilr_mvpa_others, 
                ilr_sb_lpa_sleep, 
                ilr_lpa_sleep) %>% 
  estimate_profiles(6,
                    variances = "varying",
                    covariances = "zero")

lpa_data_ilr_model2_6classes <- get_data(ilr_model2_6classes) %>%
  rename_all(tolower) %>% 
  mutate(across(class, as_factor))

ids <- ilrs_wins %>% 
  dplyr::select("id")

lpa_data_ilr_model2_6classes <- bind_cols(ids, lpa_data_ilr_model2_6classes)


# Participants N within each profile and plotting

kable(print(printToggle = FALSE, CreateTableOne(data = lpa_data_ilr_model2_6classes, c("class")))) 

# Distribution looks ok

plot_profiles(ilr_model2_6classes)

```
Note: Too few participants in class 2 and 5

### Checking the model (ilr) with 5 classes
```{r }

ilr_model2_5classes <- ilrs_wins %>% 
  dplyr::select(ilr_mvpa_others, 
                ilr_sb_lpa_sleep, 
                ilr_lpa_sleep) %>% 
  estimate_profiles(5,
                    variances = "varying",
                    covariances = "zero")

lpa_data_ilr_model2_5classes <- get_data(ilr_model2_5classes) %>%
  rename_all(tolower) %>% 
  mutate(across(class, as_factor))

ids <- ilrs_wins %>% 
  dplyr::select("id")

lpa_data_ilr_model2_5classes <- bind_cols(ids, lpa_data_ilr_model2_5classes)


# Participants N within each profile and plotting

kable(print(printToggle = FALSE, CreateTableOne(data = lpa_data_ilr_model2_5classes, c("class"))))  # Too few participants in 2 profiles

# Distribution looks ok

plot_profiles(ilr_model2_5classes)

```
Note: Too few participants in class 1 and 5

### Checking the model (ilr) with 4 classes
```{r }

ilr_model2_4classes <- ilrs_wins %>% 
  dplyr::select(ilr_mvpa_others, 
                ilr_sb_lpa_sleep, 
                ilr_lpa_sleep) %>% 
  estimate_profiles(4,
                    variances = "varying",
                    covariances = "zero")

lpa_data_ilr_model2_4classes <- get_data(ilr_model2_4classes) %>%
  rename_all(tolower) %>% 
  mutate(across(class, as_factor)) %>% 
  bind_cols(ids) %>% 
  relocate(id)

# Participants N within each profile and plotting

kable(print(printToggle = FALSE, CreateTableOne(data = lpa_data_ilr_model2_4classes, c("class")))) # Ptps are well divided between profiles

plot_profiles(ilr_model2_4classes)

```

### Checking the model (ilr) with 3 classes
```{r }

ilr_model2_3classes <- ilrs_wins %>% 
  dplyr::select(ilr_mvpa_others, 
                ilr_sb_lpa_sleep, 
                ilr_lpa_sleep) %>% 
  estimate_profiles(3,
                    variances = "varying",
                    covariances = "zero")

lpa_data_ilr_model2_3classes <- get_data(ilr_model2_3classes) %>%
  rename_all(tolower) %>% 
  mutate(across(class, as_factor)) %>% 
  bind_cols(ids) %>% 
  relocate(id)

# Participants N within each profile and plotting

kable(print(printToggle = FALSE, CreateTableOne(data = lpa_data_ilr_model2_3classes, c("class"))))

# Distribution looks ok

plot_profiles(ilr_model2_3classes)

```

### Checking the model (ilr) with 2 classes
```{r }

ilr_model2_2classes <- ilrs_wins %>% 
  dplyr::select(ilr_mvpa_others, 
                ilr_sb_lpa_sleep, 
                ilr_lpa_sleep) %>% 
  estimate_profiles(2,
                    variances = "varying",
                    covariances = "zero")

lpa_data_ilr_model2_2classes <- get_data(ilr_model2_2classes) %>%
  rename_all(tolower) %>% 
  mutate(across(class, as_factor)) %>% 
  bind_cols(ids) %>% 
  relocate(id)

# Participants N within each profile and plotting

kable(print(printToggle = FALSE, CreateTableOne(data = lpa_data_ilr_model2_2classes, c("class"))))

# Distribution looks ok

plot_profiles(ilr_model2_2classes)

```
Note: Will retain model with 4 profiles because has >25 in all profiles and has the best fit model (tbl_lpa_to_present above).

## Final LPA model and plot
```{r }

lpa_final <- lpa_data_ilr_model2_4classes
suppressMessages(lpa_final_fit <- estimate_profiles(df = lpa_data_ilr_model2_4classes[,4:6],n_profiles = 4, models = 2))
skim(lpa_final)

dat_to_plot <- reshape2::melt(lpa_data_ilr_model2_4classes[,c('class','ilr_mvpa_others', 'ilr_sb_lpa_sleep', 'ilr_lpa_sleep')],id.vars = 'class')
levels(dat_to_plot$variable) <- c('ilr 1\nMVPA: SB+LPA+Sleep', 'ilr 2\nSB: LPA+Sleep', 'ilr 3\nLPA: Sleep')

ggplot(data = dat_to_plot, aes(x = variable, y = value, fill = class)) +
  geom_boxplot() +
  scale_y_continuous(breaks = seq(-3.5,5,by = 0.2)) +
  labs(x = ' ', y = 'Isometric log-ratio') + 
  theme_light() +
  theme(legend.title = element_blank())

```

## Descriptives of ilrs
```{r,include=TRUE}
tbl_comp_mean <- CreateTableOne(data = lpa_data_ilr_model2_4classes, vars = c('ilr_mvpa_others', 'ilr_sb_lpa_sleep', 'ilr_lpa_sleep')) %>% 
  print(., printToggle = FALSE)

rownames(tbl_comp_mean)[2:4] <- c('ilr 1 (MVPA: SB+LPA+Sleep), mean (SD)',
                                  'ilr 2 (SB: LPA+Sleep), mean (SD)',
                                  'ilr 3 (LPA: Sleep), mean (SD)')

kable(tbl_comp_mean, align = 'c', vline = '',
      booktabs = TRUE,
      caption = "Descriptive statistics of ilr compositions",
      linesep = '')
```

## Descriptives of time in 24HAC as a function of LPA profile
```{r,include=TRUE}

lpa_4classes <- lpa_data_ilr_model2_4classes %>% 
  dplyr::select(1,11)

final_data_lpa <- left_join(ilrs_wins, lpa_4classes, by = "id")
skim(final_data_lpa)

final_data_lpa <- final_data_lpa %>% 
  mutate(class = case_when(class == 1 ~ "Active Chillers",
                           class == 2 ~ "Sedentary savants",
                           class == 3 ~ "Physical activity masters",
                           class == 4 ~ "Average 24HAC")) 
final_data_lpa <- final_data_lpa %>% 
  mutate(class = as.factor(class))

final_data_lpa$class <- factor(final_data_lpa$class, levels = c("Average 24HAC", "Active Chillers", "Physical activity masters", "Sedentary savants"))

# Descriptives by class
kable(print(printToggle = FALSE, CreateTableOne(data = final_data_lpa, 
                                                c("sleep_min_clos", 
                                                  "sedentary_min_clos", 
                                                  "light_pa_min_clos",
                                                  "mvpa_min_clos",
                                                  "sleep_pct_clos", 
                                                  "sedentary_pct_clos", 
                                                  "light_pa_pct_clos",
                                                  "mvpa_pct_clos",
                                                  "adas_cog_plus", 
                                                  "sex","age", "bmi", 
                                                  "bmi_cat", "moca", 
                                                  "mmse", "education"), strata = "class", 
                                                  includeNA = TRUE, addOverall = TRUE)))

# Descriptives by RCT
kable(print(printToggle = FALSE, CreateTableOne(data = final_data_lpa, 
                                                c("age","sex","bmi",
                                                  "moca","mmse", "education",
                                                  "sleep_min_clos",
                                                  "sleep_pct_clos",
                                                  "sedentary_min_clos",
                                                  "sedentary_pct _clos",
                                                  "light_pa_min_clos",
                                                  "light_pa_pct_clos",
                                                  "mvpa_min_clos", 
                                                  "mvpa_pct_clos",
                                                  "adas_cog_plus"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE)))



# Plot of classes by min and pct instead of ilrs


plot_pa_by_lpa_wrapped <- final_data_lpa %>% 
  pivot_longer(cols = 32:35, names_to = "measures", values_to = "data") %>% 
  mutate(measures = factor(measures, levels = c("sleep_min_clos",
                                                "sedentary_min_clos", 
                                                "light_pa_min_clos",
                                                "mvpa_min_clos"))) %>% # Re-ordering the x-axis variables
  ggplot(aes(y = data, x = as_factor(measures), fill = as_factor(class))) +
  geom_point(aes(colour = as_factor(class)), alpha = 0.2, position = position_jitterdodge()) +
  geom_boxplot(aes(), outlier.colour = NA, alpha = 0.5) +
  scale_x_discrete(limits = c("sleep_min_clos",
                              "sedentary_min_clos", 
                              "light_pa_min_clos",
                              "mvpa_min_clos"),
                   labels = c("Sleep", 
                              "Sedentary\nbehaviour", 
                              "Light physical\nactivity",
                              "Moderate-to-vigorous\nphysical activity")) + # Renaming labels
  scale_y_continuous(breaks=seq(0,1400,50),
                     sec.axis = sec_axis(~./1400, name = "% of a day", breaks = seq(0,1,by=0.1), labels = scales::percent)) +
  facet_wrap(~class, ncol = 4) + # Adding wrap and titles
  theme_minimal() +
  labs(y = "min/day") +
  theme(axis.text.x = element_text(angle = 45, size = 12), # Rotate x-axis label
        legend.position = "none",
        strip.text = element_text(size = 14),
        axis.title.x=element_blank()) # Font size facets titles

plot_pa_by_lpa_wrapped

# ggsave("plot_pa_lpa_20230517.png", plot_pa_by_lpa_wrapped, width = 18, height = 12, dpi = 300)

```

# Summary stats by Profile and Histograms
```{r }

kable(print(printToggle = FALSE, CreateTableOne(data = final_data_lpa, 
                                                c("sleep_min_clos", 
                                                  "sedentary_min_clos", 
                                                  "light_pa_min_clos",
                                                  "mvpa_min_clos",
                                                  "sleep_pct_clos", 
                                                  "sedentary_pct_clos", 
                                                  "light_pa_pct_clos",
                                                  "mvpa_pct_clos",
                                                  "adas_cog_plus", 
                                                  "sex","age", "bmi", 
                                                  "moca", 
                                                  "mmse", "education"), strata = "class", 
                                                  includeNA = TRUE, addOverall = TRUE)))

# Two proportionns z-test for sex
## Profile Avg vs. Active Chillers
sex_avg_ac <- final_data_lpa %>% 
  filter(class == "Average 24HAC" | class == "Active Chillers") %>% 
  dplyr::select(2,43)

table(sex_avg_ac$sex, sex_avg_ac$class)

zprop_sex_ac_sav <- prop.test(x = c(63, 43), n = c(91, 75))
zprop_sex_ac_sav # NS

## Profile Avg x PA masters
sex_avg_pam <- final_data_lpa %>% 
  filter(class == "Average 24HAC" | class == "Physical activity masters") %>% 
  dplyr::select(2,43)

table(sex_avg_pam$sex, sex_avg_pam$class)

zprop_sex_ac_pam <- prop.test(x = c(63, 35), n = c(91, 48))
zprop_sex_ac_pam # NS

## Profile Avg x SS
sex_avg_ss <- final_data_lpa %>% 
  filter(class == "Average 24HAC" | class == "Sedentary savants") %>% 
  dplyr::select(2,43)

table(sex_avg_ss$sex, sex_avg_ss$class)

sex_avg_ss <- prop.test(x = c(63, 8), n = c(91, 26))
sex_avg_ss # Sig p = 0.000922

## Profile Active Chillers x PA masters
sex_ac_pam <- final_data_lpa %>% 
  filter(class == "Active Chillers" | class == "Physical activity masters") %>% 
  dplyr::select(2,43)

table(sex_ac_pam$sex, sex_ac_pam$class)

zprop_sex_ac_pam <- prop.test(x = c(43, 35), n = c(75, 48))
zprop_sex_ac_pam # NS

## Profile Active Chillers x SS
sex_ac_ss <- final_data_lpa %>% 
  filter(class == "Active Chillers" | class == "Sedentary savants") %>% 
  dplyr::select(2,43)

table(sex_ac_ss$sex, sex_ac_ss$class)

zprop_sex_ac_ss <- prop.test(x = c(43, 8), n = c(75, 26))
zprop_sex_ac_ss # Sig 0.03512

## Profile PA masters x SS
sex_pam_ss <- final_data_lpa %>% 
  filter(class == "Physical activity masters" | class == "Sedentary savants") %>% 
  select(2,43)

table(sex_pam_ss$sex, sex_pam_ss$class)

zprop_sex_pam_ss <- prop.test(x = c(35, 8), n = c(48, 26))
zprop_sex_pam_ss # Sig 0.001109

kable(print(printToggle = FALSE, CreateTableOne(data = final_data_lpa, 
                                                c("sleep_min_clos", 
                                                  "sedentary_min_clos", 
                                                  "light_pa_min_clos",
                                                  "mvpa_min_clos",
                                                  "sleep_pct_clos", 
                                                  "sedentary_pct_clos", 
                                                  "light_pa_pct_clos",
                                                  "mvpa_pct_clos",
                                                  "adas_cog_plus", 
                                                  "sex","age", "bmi", 
                                                  "moca", 
                                                  "mmse", "education"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE)))

```

## Pairwise descriptive differences between profiles
```{r, message=FALSE}

## Select variable names for models
demo_pairw_outcomes <- c("age",
                         "bmi",
                         "sleep_min_clos", 
                         "sedentary_min_clos", 
                          "light_pa_min_clos",
                          "mvpa_min_clos",
                          "sleep_pct_clos", 
                          "sedentary_pct_clos", 
                          "light_pa_pct_clos",
                          "mvpa_pct_clos")

# List to store models
anovas_pairw_outcomes = list()

# Run models in loop script
for (var in demo_pairw_outcomes){
  # Formula
  form <- paste(
    var, "~", "class")
  anovas_pairw_outcomes[var] <- list(lm(form,
                                      data = final_data_lpa))
}

## Printing models and assumptions

for (model in anovas_pairw_outcomes){
  message("####################################################################################")
  message()
  print(formula(model), showEnv = FALSE)
  print(Anova(model, type = "III"))
  message()
  print("Post-hoc")
  print(emmeans(model, ~class))
  print(pairs(emmeans(model, ~ class)))
}


```

# ANCOVA
```{r}

adas_cog_lm <- lm(adas_cog_plus~class + sex + age + bmi + moca + rct, final_data_lpa)
Anova(adas_cog_lm, type = "III")
summary(adas_cog_lm)
performance(adas_cog_lm)
emmeans(adas_cog_lm, ~ class)
pairs(emmeans(adas_cog_lm, ~ class, adjust = "bonferroni"))

# Assumptions
## Function to summarize assumptions
assumptions <- function(x) {
  my_list <- list()
  model.metrics <- augment(x) %>% dplyr::select(-.hat, -.sigma, -.fitted)
  my_list$observations <- model.metrics %>%  count(class)
  my_list$shapiro <- model.metrics %>% shapiro_test(.resid)
  my_list$levene <- levene_test(.resid ~ class, data = model.metrics)
  my_list$outliers <- model.metrics %>% filter(abs(.std.resid) > 3)
  return(my_list)
}
assumptions(adas_cog_lm) # OK
check_model(adas_cog_lm) # OK

```
Note: NS.