---
title: "The association of the 24hour activity cycle profiles with cognitive function in older adults with Mild Cognitive Impairment: A cross-sectional study"
author: "Guilherme Moraes Balbim"
date: "01/26/2024"
output:
  pdf_document:
    latex_engine: xelatex
    toc: yes
    number_sections: yes
    toc_depth: 5
  word_document:
    toc: yes
    toc_depth: '5'
  editor_options:
    chunk_output_type: console
  chunk_output_type: console
header-includes:
- \usepackage{caption}
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
geometry: left = 1cm, right = 1cm, top = 1cm, bottom = 1cm
---

\newpage
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  fig.height = 7, fig.width = 9, fig.align = "center")
```

\small
# Loading packages
```{r message=FALSE}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse,dplyr,tidyr,lubridate,knitr,openxlsx,readxl,tidyLPA,mclust,tableone,tinytex,compositions,mgcv,skimr,GGally,rstatix,broom,psych,performance,janitor,ggplot2, PerformanceAnalytics, emmeans)

```

# Loading data for initial cleaning and descriptives
```{r message=FALSE}

data_of_interest <- read.xlsx("spreadsheets/all_data_merged_averages_20231114.xlsx")

ids_flags <- read.xlsx("spreadsheets/added_ids_flagged_20231115.xlsx")
```

# Descriptives
## Sleep, PA, SB 
```{r}
# Create a study variable
data_of_interest <- data_of_interest %>% 
  mutate(rct = str_extract(id, "[:alpha:]")) %>% 
  left_join(., ids_flags, by = "id") %>% 
  relocate(rct, .after = "id") %>% 
  relocate(new_id_flag, .after = "id")

# Exclude participants with <5 days of data, < 4h, and > 12h of sleep = 253 participants

data_of_interest <- data_of_interest %>% 
  filter(total_days >= 5)  %>% 
  filter(avg_sleep_min > 240 | avg_sleep_min < 720)

skim(data_of_interest)

# Differences by RCT
kable(print(printToggle = FALSE, CreateTableOne(data = data_of_interest, 
                                                c("avg_sleep_min",
                                                  "avg_sedentary_min",
                                                  "avg_light_pa_min",
                                                  "avg_mvpa_min"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE))) # NS

# Differences between existing and new IDs
kable(print(printToggle = FALSE, CreateTableOne(data = data_of_interest, 
                                                c("avg_sleep_min",
                                                  "avg_sedentary_min",
                                                  "avg_light_pa_min",
                                                  "avg_mvpa_min",
                                                  "adas_cog_plus"), strata = "new_id_flag", 
                                                  includeNA = TRUE, addOverall = TRUE)))

## Histograms
data_of_interest %>%
  dplyr::select("avg_sleep_min","avg_sedentary_min","avg_light_pa_min","avg_mvpa_min","avg_sleep_pct","avg_sedentary_pct","avg_light_pa_pct","avg_mvpa_pct") %>%
  pivot_longer(cols = 1:8, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  facet_wrap("measures", scale = "free") +
  geom_histogram(colour = "black", binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3))) + # Proportional bins
  theme_minimal() +
  theme(legend.position = "none")


## Bloxplot
data_of_interest %>%
  dplyr::select("avg_sleep_min","avg_sedentary_min","avg_light_pa_min","avg_mvpa_min","avg_sleep_pct","avg_sedentary_pct","avg_light_pa_pct","avg_mvpa_pct") %>%
  pivot_longer(cols = 1:8, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  geom_boxplot(colour = "black") + # Proportional bins
  ylim(-2,2) +
  facet_wrap("measures", scale = "free", ncol = 2) +
  coord_flip() +
  theme_minimal() +
  theme(legend.position = "none") # Some potential outliers
```

## Demographics
```{r}
# Continuous data
## Differences by RCT
kable(print(printToggle = FALSE, CreateTableOne(data = data_of_interest, 
                                                c("age",
                                                  "moca",
                                                  "mmse",
                                                  "height",
                                                  "weight",
                                                  "bmi"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE)))

## Differences by existing and new IDs
kable(print(printToggle = FALSE, CreateTableOne(data = data_of_interest, 
                                                c("age",
                                                  "moca",
                                                  "mmse",
                                                  "height",
                                                  "weight",
                                                  "bmi"), strata = "new_id_flag", 
                                                  includeNA = TRUE, addOverall = TRUE)))

# Categorical data
## Differences by RCT
kable(print(printToggle = FALSE, CreateTableOne(data = data_of_interest, 
                                                c("sex", "education"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE)))

## Differences by existing and new IDs
kable(print(printToggle = FALSE, CreateTableOne(data = data_of_interest, 
                                                c("sex", "education"), strata = "new_id_flag", 
                                                  includeNA = TRUE, addOverall = TRUE)))

## Normality tests
kable(digits = 3, caption = "Normality test", data_of_interest %>% 
        shapiro_test(avg_sleep_min,
                     avg_sedentary_min,
                     avg_light_pa_min,
                     avg_mvpa_min,
                     avg_sleep_pct,
                     avg_sedentary_pct,
                     avg_light_pa_pct,
                     avg_mvpa_pct)) # MVPA non-normal
```

## Outliers
```{r}
## Checking outliers
avg_sleep_min_outliers <- check_outliers(data_of_interest$avg_sleep_min,  method = "iqr")
avg_sleep_min_outliers # 4 outliers detected: cases 213, 223, 231, 245.

avg_sedentary_min_outliers <- check_outliers(data_of_interest$avg_sedentary_min,  method = "iqr")
avg_sedentary_min_outliers # 2 outliers detected: cases 40, 97.

avg_light_pa_min_outliers <- check_outliers(data_of_interest$avg_light_pa_min,  method = "iqr")
avg_light_pa_min_outliers # No outliers detected.

avg_mvpa_min_outliers <- check_outliers(data_of_interest$avg_mvpa_min, method = "iqr")
avg_mvpa_min_outliers # 11 outliers detected: cases 40, 61, 82, 93, 97, 137, 188, 220, 236, 241, 242.

# Winsorizing outliers
timeusebehav_wins <- data_of_interest

for(var in c("avg_sleep_min","avg_sedentary_min","avg_light_pa_min","avg_mvpa_min")){
  Q <- quantile(timeusebehav_wins[[var]], probs=c(.05, .95), na.rm = TRUE)
  Q1 <- quantile(timeusebehav_wins[[var]], probs=c(.25), na.rm = TRUE)
  Q3 <- quantile(timeusebehav_wins[[var]], probs=c(.75), na.rm = TRUE)
  iqr <- IQR(timeusebehav_wins[[var]], na.rm = TRUE)
  up <-  Q3+1.5*iqr # Upper Range  
  low <- Q1-1.5*iqr # Lower Range
  new <- timeusebehav_wins %>% 
    mutate(wins = ifelse(timeusebehav_wins[[var]] <= (Q1 - 1.5*iqr), Q[1], 
                         ifelse(timeusebehav_wins[[var]] >= (Q3 + 1.5*iqr), Q[2], timeusebehav_wins[[var]]))) %>% 
    select(wins) %>% 
    rename_with(~paste0(var, "_",.))
  timeusebehav_wins <- cbind(timeusebehav_wins, new)
}

## Checking distribution with winsorized data
### Histograms
timeusebehav_wins %>%
  select("avg_sleep_min_wins","avg_sedentary_min_wins","avg_light_pa_min_wins","avg_mvpa_min_wins") %>%
  pivot_longer(cols = 1:4, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  facet_wrap("measures", scale = "free") +
  geom_histogram(colour = "black", binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3))) + # Proportional bins
  theme_minimal() +
  theme(legend.position = "none")


## 24HAC behaviours by study
kable(print(printToggle = FALSE, CreateTableOne(data = timeusebehav_wins, 
                                                c("avg_sleep_min_wins",
                                                  "avg_sedentary_min_wins",
                                                  "avg_light_pa_min_wins",
                                                  "avg_mvpa_min_wins"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE))) # NS


## 24HAC behaviours by existing and new IDs
kable(print(printToggle = FALSE, CreateTableOne(data = timeusebehav_wins, 
                                                c("avg_sleep_min_wins",
                                                  "avg_sedentary_min_wins",
                                                  "avg_light_pa_min_wins",
                                                  "avg_mvpa_min_wins"), strata = "new_id_flag", 
                                                  includeNA = TRUE, addOverall = TRUE)))

data_of_interest <- timeusebehav_wins

```

## ADAS-Cog
```{r}

kable(print(printToggle = FALSE, CreateTableOne(data = data_of_interest, 
                                                c("adas_cog_plus"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE)))

## Histograms
data_of_interest %>%
  select(c("adas_cog_plus")) %>%
  pivot_longer(cols = 1:1, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  facet_wrap("measures", scale = "free") +
  geom_histogram(colour = "black", binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3))) + # Proportional bins
  theme_minimal() +
  theme(legend.position = "none")


## Bloxplot
data_of_interest %>%
  select(c("adas_cog_plus")) %>%
  pivot_longer(cols = 1, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  geom_boxplot(colour = "black") + # Proportional bins
  ylim(-2,2) +
  facet_wrap("measures", scale = "free", ncol = 2) +
  coord_flip() +
  theme_minimal() +
  theme(legend.position = "none") # Some potential outliers


# Checking outliers 
adas_cog_plus_outliers <- check_outliers(data_of_interest$adas_cog_plus, method = "iqr")
adas_cog_plus_outliers # 4 outliers detected: cases 92, 118, 139, 224.

# Winsorizing outliers
data_of_interest_wins <- data_of_interest

for(var in c("adas_cog_plus")){
  Q <- quantile(data_of_interest_wins[[var]], probs=c(.05, .95), na.rm = TRUE)
  Q1 <- quantile(data_of_interest_wins[[var]], probs=c(.25), na.rm = TRUE)
  Q3 <- quantile(data_of_interest_wins[[var]], probs=c(.75), na.rm = TRUE)
  iqr <- IQR(data_of_interest_wins[[var]], na.rm = TRUE)
  up <-  Q3+1.5*iqr # Upper Range  
  low <- Q1-1.5*iqr # Lower Range
  new <- data_of_interest_wins %>% 
    mutate(wins = ifelse(data_of_interest_wins[[var]] <= (Q1 - 1.5*iqr), Q[1], 
                         ifelse(data_of_interest_wins[[var]] >= (Q3 + 1.5*iqr), Q[2], data_of_interest_wins[[var]]))) %>% 
    select(wins) %>% 
    rename_with(~paste0(var, "_",.))
  data_of_interest_wins <- cbind(data_of_interest_wins, new)
}

## Checking distribution with winsorized data
### Histograms
data_of_interest_wins %>%
  select("adas_cog_plus_wins") %>%
  pivot_longer(cols = 1:1, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  facet_wrap("measures", scale = "free") +
  geom_histogram(colour = "black", binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3))) + # Proportional bins
  theme_minimal() +
  theme(legend.position = "none")

## Checking normality of different types of transformations for non-normal
data_of_interest_wins %>%
  select(c("adas_cog_plus_wins")) %>%
  pivot_longer(cols = 1:1, names_to = "measures", values_to = "data") %>%
  mutate(data_log = log10(data),
         data_sqrt = sqrt(data)) %>% 
  group_by(measures) %>% 
  shapiro_test(data_log,
               data_sqrt,
               data) %>% 
  kable(digits = 3) # ADAS-Cog without transformation looks better

## Checking ADAS-Cog by study
kable(print(printToggle = FALSE, CreateTableOne(data = data_of_interest_wins, 
                                                c("adas_cog_plus_wins"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE)))

# Saving the dataset 
# write.csv(data_of_interest_wins, "spreadsheets/all_data_merged_clean_wins_20240117.csv")

```

# Loading data and preparing dataset for compositions
```{r message=FALSE}

final_data <- read_csv("spreadsheets/all_data_merged_clean_wins_20240117.csv", 
                       show_col_types = FALSE,
                       name_repair = make_clean_names,
                       col_select = -c(13,15,16,20:26,28:48))
skim(final_data)

```

## Check time-use data adds up 1440
```{r message=FALSE}

# Raw
rowSums(final_data[, 
                  c("avg_mvpa_min", "avg_sedentary_min", "avg_light_pa_min", "avg_sleep_min")
]) # No, will need the closure from compositions
final_data <- final_data %>% 
  mutate(day_duration_wins = rowSums(final_data[,c("avg_mvpa_min_wins", "avg_sedentary_min_wins", "avg_light_pa_min_wins", "avg_sleep_min_wins")]),
         day_duration = rowSums(final_data[,c("avg_mvpa_min", "avg_sedentary_min", "avg_light_pa_min", "avg_sleep_min")]))

# Winsorized
rowSums(final_data[, 
                  c("avg_mvpa_min_wins", "avg_sedentary_min_wins", "avg_light_pa_min_wins", "avg_sleep_min_wins")
]) # No, will need the closure from compositions
  
```

## Closure function (compositions package)
```{r message=FALSE}

# Winsorized
## Create dataframes containing the outcome, predictors and covariates
cols_outcome_wins <- c("adas_cog_plus_wins")
cols_pred_min_wins <- c("avg_mvpa_min_wins", "avg_sedentary_min_wins", "avg_light_pa_min_wins", "avg_sleep_min_wins")
cols_id_covars <- c("id","sex","age", "bmi", "moca", "mmse", "education", "rct")

## Combine all dataframes
cols_pairwise_wins <- c(cols_outcome_wins, cols_pred_min_wins, cols_id_covars)

## Create new dataframe containing all rows from final_data but only cols_pairwise_wins columns
pairwise_data_wins <- final_data[, cols_pairwise_wins] %>% 
  relocate(id)

## Ensuring that time-use data have a total of 1440min
behav_min_wins <- pairwise_data_wins %>% 
  select("avg_mvpa_min_wins", "avg_sedentary_min_wins", "avg_light_pa_min_wins", "avg_sleep_min_wins")

behav_min_wins_clo <- data.frame(clo(behav_min_wins, total = 1440)) %>% 
    rename_with(~gsub("wins", "wins_clos", .), contains("wins"))

rowSums(behav_min_wins_clo) # All 1440min

# Raw
## Create dataframes containing the outcome, predictors and covariates
cols_outcome <- c("adas_cog_plus")
cols_pred_min <- c("avg_mvpa_min", "avg_sedentary_min", "avg_light_pa_min", "avg_sleep_min")
cols_id_covars <- c("id","sex","age", "bmi", "moca", "mmse", "education", "rct")

## Combine all dataframes
cols_pairwise <- c(cols_outcome, cols_pred_min, cols_id_covars)

## Create new dataframe containing all rows from final_data but only cols_pairwise columns
pairwise_data <- final_data[, cols_pairwise] %>% 
  relocate(id)

## Ensuring that time-use data have a total of 1440min
behav_min <- pairwise_data %>% 
  select("avg_mvpa_min", "avg_sedentary_min", "avg_light_pa_min", "avg_sleep_min")

behav_min_clo <- data.frame(clo(behav_min, total = 1440)) %>% 
    rename_with(~gsub("min", "min_clos", .), contains("min"))

rowSums(behav_min_clo) # All 1440min

# Appending 24HAC winsorized and raw closure behaviours to final data
behav_min_wins_clo_id <- cbind(final_data$id, behav_min_wins_clo) %>% 
  rename("id" = "final_data$id") %>% 
  bind_cols(., behav_min_clo)

all_data <- left_join(final_data, behav_min_wins_clo_id, by = "id")

```

## Creating percentage from closure data
```{r message=FALSE}

all_data <- all_data %>% 
  mutate(avg_mvpa_pct_clos = (avg_mvpa_min_clos/1440)*100,
         avg_light_pa_pct_clos = (avg_light_pa_min_clos/1440)*100,
         avg_sedentary_pct_clos = (avg_sedentary_min_clos/1440)*100,
         avg_sleep_pct_clos = (avg_sleep_min_clos/1440)*100,
         avg_mvpa_pct_wins_clos = (avg_mvpa_min_wins_clos/1440)*100,
         avg_light_pa_pct_wins_clos = (avg_light_pa_min_wins_clos/1440)*100,
         avg_sedentary_pct_wins_clos = (avg_sedentary_min_wins_clos/1440)*100,
         avg_sleep_pct_wins_clos = (avg_sleep_min_wins_clos/1440)*100)

```

# Compositional data analysis
## 3 ilrs for LPA
Setting up the data to create the ilrs so that the first ilrs ilr1 = MVPA:all remaining, ilr2 = SB:LPA+sleep, ilr3 = LPA:sleep. All three ilrs will be entered in to LPA analysis to capture the entire 24-hour time-use composition.
```{r,include=TRUE}

## Create a matrix
data_comp_mean_min <- matrix(c(all_data$avg_mvpa_min_wins_clos, 
                               all_data$avg_sedentary_min_wins_clos, 
                               all_data$avg_light_pa_min_wins_clos, 
                               all_data$avg_sleep_min_wins_clos), ncol = 4)

data_min_clos <- all_data %>% 
  select("avg_mvpa_min_wins_clos", "avg_sedentary_min_wins_clos", 
         "avg_light_pa_min_wins_clos", "avg_sleep_min_wins_clos")

rowSums(data_min_clos)

data_comp_mean_pct <- matrix(c(all_data$avg_mvpa_pct_wins_clos,
                               all_data$avg_sedentary_pct_wins_clos,
                               all_data$avg_light_pa_pct_wins_clos,
                               all_data$avg_sleep_pct_wins_clos), ncol = 4)

data_pct_clos <- all_data %>% 
  select("avg_mvpa_pct_wins_clos", "avg_sedentary_pct_wins_clos", 
         "avg_light_pa_pct_wins_clos", "avg_sleep_pct_wins_clos")

rowSums(data_pct_clos)

## ILR coordinates - sequential binary partition process
sbp4_3 = matrix(c( 1, -1, -1,-1, 
                 0, 1, -1, -1,
                 0, 0, 1, -1),
              ncol=4, byrow=TRUE)


psi4_3 = gsi.buildilrBase(t(sbp4_3))

## With 3 ilrs. Obs: "avg_mvpa_min_wins_clos", "avg_sedentary_min_wins_clos", "avg_light_pa_min_wins_clos", "avg_sleep_min_wins_clos"

ilrs3 = ilr(acomp(all_data[, c("avg_mvpa_pct_wins_clos", "avg_sedentary_pct_wins_clos", 
         "avg_light_pa_pct_wins_clos", "avg_sleep_pct_wins_clos")]), V=psi4_3) # compute ilrs of all rows but only columns in cols_pred2 (with closure)
head(ilrs3)

# rename ilrs to ilr1, ilr2 and ilr3
colnames(ilrs3) <- paste0("ilr", 1:3)
head(ilrs3)
nrow(ilrs3)
nrow(all_data) # Number of rows for ilrs and all_data_comp matches

# Get ilrs means
ilr_mean_min <- colMeans(ilrs3)
ilr_mean_min

```

# Descriptives of 24HAC behaviours (compositional means of whole population)
```{r,include=TRUE}
tbl1 <- CreateTableOne(data = all_data, vars = c('sex',
                                                   'avg_sleep_min_wins_clos',
                                                   'avg_sleep_pct_wins_clos',
                                                   'avg_sedentary_min_wins_clos',
                                                   'avg_sedentary_pct_wins_clos',
                                                   'avg_light_pa_min_wins_clos',
                                                   'avg_light_pa_pct_wins_clos',
                                                   'avg_mvpa_min_wins_clos',
                                                   'avg_mvpa_pct_wins_clos')) %>% 
  print(., printToggle = FALSE)

tbl1  
  
rownames(tbl1)[2:10] <- c('Sex = Male (%)',
                          'Sleep (min/day), mean (SD)',
                         'Sleep (%/day), mean (SD)',
                         'Sedentary behaviour (min/day), mean (SD)',
                         'Sedentary behaviour (%/day), mean (SD)',
                         'Light PA (min/day), mean (SD)',
                         'Light PA (%/day), mean (SD)',
                         'Moderate-to-vigorous PA (min/day), mean (SD)',
                         'Moderate-to-vigorous PA (%/day), mean (SD)')

kable(tbl1, align = 'c', vline = '',
      booktabs = TRUE,
      caption = "Descriptive statistics of the sample",
      linesep = '')

```

## Create new dataframe containing the data to be used in LPA
```{r message=FALSE}

ilr3_data <- cbind(all_data, ilrs3) %>% 
  rename_all( ~ str_replace(., "avg_", "")) %>% 
  rename("ilr_mvpa_others" = "ilr1",
         "ilr_sb_lpa_sleep" = "ilr2",
         "ilr_lpa_sleep" = "ilr3")
            
# Double-check there are no NAs in the data
table(is.na(ilr3_data)) # No NAs
skim(ilr3_data)

# Save dataset
#write.csv(ilr3_data, "spreadsheets/all_data_3ilrs_20240117.csv")

```

# Latent Profile Analysis with 3 ilrs
## Checking normality and dealing with outliers
```{r message=FALSE}

## Histograms
ilr3_data %>%
  select(c("ilr_mvpa_others", "ilr_sb_lpa_sleep", "ilr_lpa_sleep")) %>%
  pivot_longer(cols = 1:3, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  facet_wrap("measures", scale = "free") +
  geom_histogram(colour = "black", binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3))) + # Proportional bins
  theme_minimal() +
  theme(legend.position = "none")

# Checking outliers 
ilr1_outliers <- check_outliers(ilr3_data$ilr_mvpa_others, method = "iqr")
ilr1_outliers # 4 outliers detected: cases 35, 197, 223, 239.

ilr2_outliers <- check_outliers(ilr3_data$ilr_sb_lpa_sleep, method = "iqr")
ilr2_outliers # No outliers detected.

ilr3_outliers <- check_outliers(ilr3_data$ilr_lpa_sleep, method = "iqr")
ilr3_outliers # 1 outlier detected: case 204.

# Winsorizing outliers
ilrs_wins <- ilr3_data

for(var in c("ilr_mvpa_others", "ilr_sb_lpa_sleep", "ilr_lpa_sleep")){
  Q <- quantile(ilrs_wins[[var]], probs=c(.05, .95), na.rm = TRUE)
  Q1 <- quantile(ilrs_wins[[var]], probs=c(.25), na.rm = TRUE)
  Q3 <- quantile(ilrs_wins[[var]], probs=c(.75), na.rm = TRUE)
  iqr <- IQR(ilrs_wins[[var]], na.rm = TRUE)
  up <-  Q3+1.5*iqr # Upper Range  
  low <- Q1-1.5*iqr # Lower Range
  new <- ilrs_wins %>% 
    mutate(wins = ifelse(ilrs_wins[[var]] <= (Q1 - 1.5*iqr), Q[1], 
                         ifelse(ilrs_wins[[var]] >= (Q3 + 1.5*iqr), Q[2], ilrs_wins[[var]]))) %>% 
    select(wins) %>% 
    rename_with(~paste0(var, "_",.))
  ilrs_wins <- cbind(ilrs_wins, new)
}

## Checking distribution with winsorized data
### Histograms
ilrs_wins %>%
  select("ilr_mvpa_others_wins", "ilr_sb_lpa_sleep_wins", "ilr_lpa_sleep_wins") %>%
  pivot_longer(cols = 1:3, names_to = "measures", values_to = "data") %>%
  ggplot(aes(data, fill = measures)) +
  facet_wrap("measures", scale = "free") +
  geom_histogram(colour = "black", binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3))) + # Proportional bins
  theme_minimal() +
  theme(legend.position = "none") # All good now

# Checking outliers 
ilr1_outliers <- check_outliers(ilrs_wins$ilr_mvpa_others_wins, method = "iqr")
ilr1_outliers # No outliers

ilr2_outliers <- check_outliers(ilrs_wins$ilr_sb_lpa_sleep_wins, method = "iqr")
ilr2_outliers # No outliers

ilr3_outliers <- check_outliers(ilrs_wins$ilr_lpa_sleep_wins, method = "iqr")
ilr3_outliers # No outliers


skim(ilrs_wins)

```

## Exploring LPAs
### Winsorized
```{r }

# Comparing solutions from 2 to 6 classes

suppressMessages(mod <- estimate_profiles(df = ilrs_wins[,44:46],n_profiles = 2:6, models = 2))
tbl_lpa <- sapply(mod,function(x) {x$fit}) %>% t(.)
tbl_lpa[,3:17] <- round(tbl_lpa[,3:17],2)
tbl_lpa_to_present <- tbl_lpa[,c(2,3,4,6,7,8,9,10,11,15,18,12)]
rownames(tbl_lpa_to_present) <- c('2-class model','3-class model','4-class model','5-class model','6-class model')
tbl_lpa_to_present

```

#### Checking the model (ilr) with 6 classes
```{r }

ilr_model2_6classes <- ilrs_wins %>% 
  dplyr::select(ilr_mvpa_others_wins, 
                ilr_sb_lpa_sleep_wins, 
                ilr_lpa_sleep_wins) %>% 
  estimate_profiles(6,
                    variances = "varying",
                    covariances = "zero")

lpa_data_ilr_model2_6classes <- get_data(ilr_model2_6classes) %>%
  rename_all(tolower) %>% 
  mutate(across(class, as_factor))

ids <- ilrs_wins %>% 
  dplyr::select("id")

lpa_data_ilr_model2_6classes <- bind_cols(ids, lpa_data_ilr_model2_6classes)


# Participants N within each profile and plotting

kable(print(printToggle = FALSE, CreateTableOne(data = lpa_data_ilr_model2_6classes, c("class")))) # One class <10

plot_profiles(ilr_model2_6classes)

```
Note: Too few participants in class 2 and 5

#### Checking the model (ilr) with 5 classes
```{r }

ilr_model2_5classes <- ilrs_wins %>% 
  dplyr::select(ilr_mvpa_others_wins, 
                ilr_sb_lpa_sleep_wins, 
                ilr_lpa_sleep_wins) %>% 
  estimate_profiles(5,
                    variances = "varying",
                    covariances = "zero")

lpa_data_ilr_model2_5classes <- get_data(ilr_model2_5classes) %>%
  rename_all(tolower) %>% 
  mutate(across(class, as_factor))

ids <- ilrs_wins %>% 
  dplyr::select("id")

lpa_data_ilr_model2_5classes <- bind_cols(ids, lpa_data_ilr_model2_5classes)


# Participants N within each profile and plotting

kable(print(printToggle = FALSE, CreateTableOne(data = lpa_data_ilr_model2_5classes, c("class"))))  

# Distribution looks ok

plot_profiles(ilr_model2_5classes)

```

#### Checking the model (ilr) with 4 classes
```{r }

ilr_model2_4classes <- ilrs_wins %>% 
  dplyr::select(ilr_mvpa_others_wins, 
                ilr_sb_lpa_sleep_wins, 
                ilr_lpa_sleep_wins) %>% 
  estimate_profiles(4,
                    variances = "varying",
                    covariances = "zero")

lpa_data_ilr_model2_4classes <- get_data(ilr_model2_4classes) %>%
  rename_all(tolower) %>% 
  mutate(across(class, as_factor)) %>% 
  bind_cols(ids) %>% 
  relocate(id)

# Participants N within each profile and plotting

kable(print(printToggle = FALSE, CreateTableOne(data = lpa_data_ilr_model2_4classes, c("class")))) # Ptps are well divided between profiles

plot_profiles(ilr_model2_4classes)

```

#### Checking the model (ilr) with 3 classes
```{r }

ilr_model2_3classes <- ilrs_wins %>% 
  dplyr::select(ilr_mvpa_others_wins, 
                ilr_sb_lpa_sleep_wins, 
                ilr_lpa_sleep_wins) %>% 
  estimate_profiles(3,
                    variances = "varying",
                    covariances = "zero")

lpa_data_ilr_model2_3classes <- get_data(ilr_model2_3classes) %>%
  rename_all(tolower) %>% 
  mutate(across(class, as_factor)) %>% 
  bind_cols(ids) %>% 
  relocate(id)

# Participants N within each profile and plotting

kable(print(printToggle = FALSE, CreateTableOne(data = lpa_data_ilr_model2_3classes, c("class"))))

# Distribution looks ok

plot_profiles(ilr_model2_3classes)

```

#### Checking the model (ilr) with 2 classes
```{r }

ilr_model2_2classes <- ilrs_wins %>% 
  dplyr::select(ilr_mvpa_others_wins, 
                ilr_sb_lpa_sleep_wins, 
                ilr_lpa_sleep_wins) %>% 
  estimate_profiles(2,
                    variances = "varying",
                    covariances = "zero")

lpa_data_ilr_model2_2classes <- get_data(ilr_model2_2classes) %>%
  rename_all(tolower) %>% 
  mutate(across(class, as_factor)) %>% 
  bind_cols(ids) %>% 
  relocate(id)

# Participants N within each profile and plotting

kable(print(printToggle = FALSE, CreateTableOne(data = lpa_data_ilr_model2_2classes, c("class"))))

# Distribution looks ok

plot_profiles(ilr_model2_2classes)

```
Note: Will retain model with 4 profiles because has >25 in all profiles and has the best fit model (tbl_lpa_to_present above).

#### Final LPA model and plot
```{r }

lpa_final <- lpa_data_ilr_model2_4classes
suppressMessages(lpa_final_fit <- estimate_profiles(df = lpa_data_ilr_model2_4classes[,4:6],n_profiles = 4, models = 2))
skim(lpa_final)

dat_to_plot <- reshape2::melt(lpa_data_ilr_model2_4classes[,c('class','ilr_mvpa_others_wins','ilr_sb_lpa_sleep_wins',
                                                              'ilr_lpa_sleep_wins')],id.vars = 'class')
levels(dat_to_plot$variable) <- c('ilr 1\nMVPA: SB+LPA+Sleep', 'ilr 2\nSB: LPA+Sleep', 'ilr 3\nLPA: Sleep')

ggplot(data = dat_to_plot, aes(x = variable, y = value, fill = class)) +
  geom_boxplot() +
  scale_y_continuous(breaks = seq(-5,5,by = 0.2)) +
  labs(x = ' ', y = 'Isometric log-ratio') + 
  theme_light() +
  theme(legend.title = element_blank())



```

#### Descriptives of ilrs
```{r,include=TRUE}
tbl_comp_mean <- CreateTableOne(data = lpa_data_ilr_model2_4classes,vars = c('ilr_mvpa_others_wins','ilr_sb_lpa_sleep_wins',
                                                                             'ilr_lpa_sleep_wins')) %>% 
  print(., printToggle = FALSE)

rownames(tbl_comp_mean)[2:4] <- c('ilr 1 (MVPA: SB+LPA+Sleep), mean (SD)',
                                  'ilr 2 (SB: LPA+Sleep), mean (SD)',
                                  'ilr 3 (LPA: Sleep), mean (SD)')

kable(tbl_comp_mean, align = 'c', vline = '',
      booktabs = TRUE,
      caption = "Descriptive statistics of ilr compositions",
      linesep = '')
```


### Raw
```{r }

# Comparing solutions from 2 to 6 classes

suppressMessages(mod <- estimate_profiles(df = ilrs_wins[,41:43],n_profiles = 2:6, models = 2))
tbl_lpa <- sapply(mod,function(x) {x$fit}) %>% t(.)
tbl_lpa[,3:17] <- round(tbl_lpa[,3:17],2)
tbl_lpa_to_present <- tbl_lpa[,c(2,3,4,6,7,8,9,10,11,15,18,12)]
rownames(tbl_lpa_to_present) <- c('2-class model','3-class model','4-class model','5-class model','6-class model')
tbl_lpa_to_present

```

#### Checking the model (ilr) with 6 classes
```{r }

ilr_model2_6classes <- ilrs_wins %>% 
  dplyr::select(ilr_mvpa_others, 
                ilr_sb_lpa_sleep, 
                ilr_lpa_sleep) %>% 
  estimate_profiles(6,
                    variances = "varying",
                    covariances = "zero")

lpa_data_ilr_model2_6classes <- get_data(ilr_model2_6classes) %>%
  rename_all(tolower) %>% 
  mutate(across(class, as_factor))

ids <- ilrs_wins %>% 
  dplyr::select("id")

lpa_data_ilr_model2_6classes <- bind_cols(ids, lpa_data_ilr_model2_6classes)


# Participants N within each profile and plotting

kable(print(printToggle = FALSE, CreateTableOne(data = lpa_data_ilr_model2_6classes, c("class")))) # One class <10

plot_profiles(ilr_model2_6classes)

```
Note: Too few participants in class 3

#### Checking the model (ilr) with 5 classes
```{r }

ilr_model2_5classes <- ilrs_wins %>% 
  dplyr::select(ilr_mvpa_others, 
                ilr_sb_lpa_sleep, 
                ilr_lpa_sleep) %>% 
  estimate_profiles(5,
                    variances = "varying",
                    covariances = "zero")

lpa_data_ilr_model2_5classes <- get_data(ilr_model2_5classes) %>%
  rename_all(tolower) %>% 
  mutate(across(class, as_factor))

ids <- ilrs_wins %>% 
  dplyr::select("id")

lpa_data_ilr_model2_5classes <- bind_cols(ids, lpa_data_ilr_model2_5classes)


# Participants N within each profile and plotting

kable(print(printToggle = FALSE, CreateTableOne(data = lpa_data_ilr_model2_5classes, c("class"))))  

# Distribution looks ok

plot_profiles(ilr_model2_5classes)

```

#### Checking the model (ilr) with 4 classes
```{r }

ilr_model2_4classes <- ilrs_wins %>% 
  dplyr::select(ilr_mvpa_others, 
                ilr_sb_lpa_sleep, 
                ilr_lpa_sleep) %>% 
  estimate_profiles(4,
                    variances = "varying",
                    covariances = "zero")

lpa_data_ilr_model2_4classes <- get_data(ilr_model2_4classes) %>%
  rename_all(tolower) %>% 
  mutate(across(class, as_factor)) %>% 
  bind_cols(ids) %>% 
  relocate(id)

# Participants N within each profile and plotting

kable(print(printToggle = FALSE, CreateTableOne(data = lpa_data_ilr_model2_4classes, c("class")))) # Ptps are well divided between profiles

plot_profiles(ilr_model2_4classes)

```

#### Checking the model (ilr) with 3 classes
```{r }

ilr_model2_3classes <- ilrs_wins %>% 
  dplyr::select(ilr_mvpa_others, 
                ilr_sb_lpa_sleep, 
                ilr_lpa_sleep) %>% 
  estimate_profiles(3,
                    variances = "varying",
                    covariances = "zero")

lpa_data_ilr_model2_3classes <- get_data(ilr_model2_3classes) %>%
  rename_all(tolower) %>% 
  mutate(across(class, as_factor)) %>% 
  bind_cols(ids) %>% 
  relocate(id)

# Participants N within each profile and plotting

kable(print(printToggle = FALSE, CreateTableOne(data = lpa_data_ilr_model2_3classes, c("class"))))

# Distribution looks ok

plot_profiles(ilr_model2_3classes)

```

#### Checking the model (ilr) with 2 classes
```{r }

ilr_model2_2classes <- ilrs_wins %>% 
  dplyr::select(ilr_mvpa_others, 
                ilr_sb_lpa_sleep, 
                ilr_lpa_sleep) %>% 
  estimate_profiles(2,
                    variances = "varying",
                    covariances = "zero")

lpa_data_ilr_model2_2classes <- get_data(ilr_model2_2classes) %>%
  rename_all(tolower) %>% 
  mutate(across(class, as_factor)) %>% 
  bind_cols(ids) %>% 
  relocate(id)

# Participants N within each profile and plotting

kable(print(printToggle = FALSE, CreateTableOne(data = lpa_data_ilr_model2_2classes, c("class"))))

# Distribution looks ok

plot_profiles(ilr_model2_2classes)

```
Note: Will retain model with 4 profiles because has >25 in all profiles and has the best fit model (tbl_lpa_to_present above).

#### Final LPA model and plot
```{r }

lpa_final <- lpa_data_ilr_model2_4classes
suppressMessages(lpa_final_fit <- estimate_profiles(df = lpa_data_ilr_model2_4classes[,4:6],n_profiles = 4, models = 2))
skim(lpa_final)

dat_to_plot <- reshape2::melt(lpa_data_ilr_model2_4classes[,c('class','ilr_mvpa_others', 'ilr_sb_lpa_sleep', 'ilr_lpa_sleep')],id.vars = 'class')
levels(dat_to_plot$variable) <- c('ilr 1\nMVPA: SB+LPA+Sleep', 'ilr 2\nSB: LPA+Sleep', 'ilr 3\nLPA: Sleep')

ggplot(data = dat_to_plot, aes(x = variable, y = value, fill = class)) +
  geom_boxplot() +
  scale_y_continuous(breaks = seq(-5,5,by = 0.2)) +
  labs(x = ' ', y = 'Isometric log-ratio') + 
  theme_light() +
  theme(legend.title = element_blank())



```

#### Descriptives of ilrs
```{r,include=TRUE}
tbl_comp_mean <- CreateTableOne(data = lpa_data_ilr_model2_4classes, vars = c('ilr_mvpa_others', 'ilr_sb_lpa_sleep', 'ilr_lpa_sleep')) %>% 
  print(., printToggle = FALSE)

rownames(tbl_comp_mean)[2:4] <- c('ilr 1 (MVPA: SB+LPA+Sleep), mean (SD)',
                                  'ilr 2 (SB: LPA+Sleep), mean (SD)',
                                  'ilr 3 (LPA: Sleep), mean (SD)')

kable(tbl_comp_mean, align = 'c', vline = '',
      booktabs = TRUE,
      caption = "Descriptive statistics of ilr compositions",
      linesep = '')
```


# Summary stats by Profile and Histograms
## Winsorized
### Pairwise descriptive difference by class
```{r }

kable(print(printToggle = FALSE, CreateTableOne(data = final_data_lpa, 
                                                c("age","sex","bmi","moca", "mmse","day_duration_wins",
                                                  "sleep_min_wins_clos", 
                                                  "sedentary_min_wins_clos", 
                                                  "light_pa_min_wins_clos",
                                                  "mvpa_min_wins_clos",
                                                  "sleep_pct_wins_clos", 
                                                  "sedentary_pct_wins_clos", 
                                                  "light_pa_pct_wins_clos",
                                                  "mvpa_pct_wins_clos",
                                                  "adas_cog_plus_wins"), strata = "class", 
                                                  includeNA = TRUE, addOverall = TRUE)))

# Two proportionns z-test for sex
## Profile Avg vs. Active Chillers
sex_avg_ac <- final_data_lpa %>% 
  filter(class == "Average 24-HAC" | class == "Active Chillers") %>% 
  dplyr::select(4,47) # Selecting sex and class

table(sex_avg_ac$sex, sex_avg_ac$class)

zprop_sex_ac_sav <- prop.test(x = c(75, 36), n = c(108, 64)) # prop.test(x = c(f, f), n = c(n, n))
zprop_sex_ac_sav # NS

## Profile Avg x PA masters
sex_avg_pam <- final_data_lpa %>% 
  filter(class == "Average 24HAC" | class == "Physical activity masters") %>% 
  dplyr::select(4,47) # Selecting sex and class

table(sex_avg_pam$sex, sex_avg_pam$class)

zprop_sex_ac_pam <- prop.test(x = c(75, 42), n = c(108, 56)) # prop.test(x = c(f, f), n = c(n, n))
zprop_sex_ac_pam # NS

## Profile Avg x SS
sex_avg_ss <- final_data_lpa %>% 
  filter(class == "Average 24HAC" | class == "Sedentary savants") %>% 
  dplyr::select(4,47) # Selecting sex and class

table(sex_avg_ss$sex, sex_avg_ss$class)

sex_avg_ss <- prop.test(x = c(75, 7), n = c(108, 25))
sex_avg_ss # Sig p = 0.0003036

## Profile Active Chillers x PA masters
sex_ac_pam <- final_data_lpa %>% 
  filter(class == "Active Chillers" | class == "Physical activity masters") %>% 
  dplyr::select(4,47) # Selecting sex and class

table(sex_ac_pam$sex, sex_ac_pam$class)

zprop_sex_ac_pam <- prop.test(x = c(33, 42), n = c(64, 56))
zprop_sex_ac_pam # NS

## Profile Active Chillers x SS
sex_ac_ss <- final_data_lpa %>% 
  filter(class == "Active Chillers" | class == "Sedentary savants") %>% 
  dplyr::select(4,47) # Selecting sex and class

table(sex_ac_ss$sex, sex_ac_ss$class)

zprop_sex_ac_ss <- prop.test(x = c(39, 7), n = c(70, 26))
zprop_sex_ac_ss # Sig 0.01402

## Profile PA masters x SS
sex_pam_ss <- final_data_lpa %>% 
  filter(class == "Physical activity masters" | class == "Sedentary savants") %>% 
  dplyr::select(4,47) # Selecting sex and class

table(sex_pam_ss$sex, sex_pam_ss$class)

zprop_sex_pam_ss <- prop.test(x = c(42, 7), n = c(56, 25))
zprop_sex_pam_ss # Sig 0.0001762

kable(print(printToggle = FALSE, CreateTableOne(data = final_data_lpa, 
                                                c("age","sex","bmi","moca","mmse","day_duration_wins","education",
                                                  "sleep_min_wins_clos",
                                                  "sleep_pct_wins_clos", 
                                                  "sedentary_min_wins_clos",
                                                  "sedentary_pct_wins_clos", 
                                                  "light_pa_min_wins_clos",
                                                  "light_pa_pct_wins_clos",
                                                  "mvpa_min_wins_clos",
                                                  "mvpa_pct_wins_clos",
                                                  "adas_cog_plus_wins"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE)))

```

### Pairwise continuous descriptive differences between profiles
```{r, message=FALSE}

## Select variable names for models
demo_pairw_outcomes <- c("age",
                         "bmi",
                         "day_duration_wins",
                         "sleep_pct_wins_clos", 
                         "sedentary_pct_wins_clos", 
                         "light_pa_pct_wins_clos",
                         "mvpa_pct_wins_clos")

# List to store models
anovas_pairw_outcomes = list()

# Run models in loop script
for (var in demo_pairw_outcomes){
  # Formula
  form <- paste(
    var, "~", "class")
  anovas_pairw_outcomes[var] <- list(lm(form,
                                      data = final_data_lpa))
}

## Printing models and assumptions

for (model in anovas_pairw_outcomes){
  message("####################################################################################")
  message()
  print(formula(model), showEnv = FALSE)
  print(Anova(model, type = "III"))
  message()
  print("Post-hoc")
  print(emmeans(model, ~class))
  print(pairs(emmeans(model, ~ class)))
}

final_data_lpa <- final_data_lpa %>%
  mutate(across(where(is.character), ~iconv(., "UTF-8", "ASCII//TRANSLIT")))

```

## Raw
### Pairwise descriptive difference by class
```{r }

kable(print(printToggle = FALSE, CreateTableOne(data = final_data_lpa, 
                                                c("age","sex","bmi","moca", "mmse","day_duration",
                                                  "sleep_min_clos", 
                                                  "sedentary_min_clos", 
                                                  "light_pa_min_clos",
                                                  "mvpa_min_clos",
                                                  "sleep_pct_clos", 
                                                  "sedentary_pct_clos", 
                                                  "light_pa_pct_clos",
                                                  "mvpa_pct_clos",
                                                  "adas_cog_plus"), strata = "class", 
                                                  includeNA = TRUE, addOverall = TRUE)))

# Two proportionns z-test for sex
## Profile Avg vs. Active Chillers
sex_avg_ac <- final_data_lpa %>% 
  filter(class == "Average 24-HAC" | class == "Active Chillers") %>% 
  dplyr::select(4,47) # Selecting sex and class

table(sex_avg_ac$sex, sex_avg_ac$class)

zprop_sex_ac_sav <- prop.test(x = c(75, 36), n = c(108, 64)) # prop.test(x = c(f, f), n = c(n, n))
zprop_sex_ac_sav # NS

## Profile Avg x PA masters
sex_avg_pam <- final_data_lpa %>% 
  filter(class == "Average 24HAC" | class == "Physical activity masters") %>% 
  dplyr::select(4,47) # Selecting sex and class

table(sex_avg_pam$sex, sex_avg_pam$class)

zprop_sex_ac_pam <- prop.test(x = c(75, 42), n = c(108, 56)) # prop.test(x = c(f, f), n = c(n, n))
zprop_sex_ac_pam # NS

## Profile Avg x SS
sex_avg_ss <- final_data_lpa %>% 
  filter(class == "Average 24HAC" | class == "Sedentary savants") %>% 
  dplyr::select(4,47) # Selecting sex and class

table(sex_avg_ss$sex, sex_avg_ss$class)

sex_avg_ss <- prop.test(x = c(75, 7), n = c(108, 25))
sex_avg_ss # Sig p = 0.0003036

## Profile Active Chillers x PA masters
sex_ac_pam <- final_data_lpa %>% 
  filter(class == "Active Chillers" | class == "Physical activity masters") %>% 
  dplyr::select(4,47) # Selecting sex and class

table(sex_ac_pam$sex, sex_ac_pam$class)

zprop_sex_ac_pam <- prop.test(x = c(33, 42), n = c(64, 56))
zprop_sex_ac_pam # NS

## Profile Active Chillers x SS
sex_ac_ss <- final_data_lpa %>% 
  filter(class == "Active Chillers" | class == "Sedentary savants") %>% 
  dplyr::select(4,47) # Selecting sex and class

table(sex_ac_ss$sex, sex_ac_ss$class)

zprop_sex_ac_ss <- prop.test(x = c(39, 7), n = c(70, 26))
zprop_sex_ac_ss # Sig 0.01402

## Profile PA masters x SS
sex_pam_ss <- final_data_lpa %>% 
  filter(class == "Physical activity masters" | class == "Sedentary savants") %>% 
  dplyr::select(4,47) # Selecting sex and class

table(sex_pam_ss$sex, sex_pam_ss$class)

zprop_sex_pam_ss <- prop.test(x = c(42, 7), n = c(56, 25))
zprop_sex_pam_ss # Sig 0.0001762

kable(print(printToggle = FALSE, CreateTableOne(data = final_data_lpa, 
                                                c("age","sex","bmi","moca", "mmse","day_duration",
                                                  "sleep_min_clos", 
                                                  "sedentary_min_clos", 
                                                  "light_pa_min_clos",
                                                  "mvpa_min_clos",
                                                  "sleep_pct_clos", 
                                                  "sedentary_pct_clos", 
                                                  "light_pa_pct_clos",
                                                  "mvpa_pct_clos",
                                                  "adas_cog_plus"), strata = "rct", 
                                                  includeNA = TRUE, addOverall = TRUE)))

```

### Pairwise continuous descriptive differences between profiles
```{r, message=FALSE}

## Select variable names for models
demo_pairw_outcomes <- c("age",
                         "bmi",
                         "moca",
                         "mmse",
                         "day_duration",
                         "sleep_pct_clos", 
                         "sedentary_pct_clos", 
                         "light_pa_pct_clos",
                         "mvpa_pct_clos")

# List to store models
anovas_pairw_outcomes = list()

# Run models in loop script
for (var in demo_pairw_outcomes){
  # Formula
  form <- paste(
    var, "~", "class")
  anovas_pairw_outcomes[var] <- list(lm(form,
                                      data = final_data_lpa))
}

## Printing models and assumptions

for (model in anovas_pairw_outcomes){
  message("####################################################################################")
  message()
  print(formula(model), showEnv = FALSE)
  print(Anova(model, type = "III"))
  message()
  print("Post-hoc")
  print(emmeans(model, ~class))
  print(pairs(emmeans(model, ~ class)))
}

final_data_lpa <- final_data_lpa %>%
  mutate(across(where(is.character), ~iconv(., "UTF-8", "ASCII//TRANSLIT")))

```

# ANCOVA on ADAS-Cog
## Winsorized
```{r}

adas_cog_lm <- lm(adas_cog_plus_wins~class + sex + age + bmi + moca + rct, final_data_lpa)
Anova(adas_cog_lm, type = "III")
summary(adas_cog_lm)
performance(adas_cog_lm)
emmeans(adas_cog_lm, ~ class)
pairs(emmeans(adas_cog_lm, ~ class, adjust = "tukey"))

# Assumptions
## Function to summarize assumptions
assumptions <- function(x) {
  my_list <- list()
  model.metrics <- augment(x) %>% dplyr::select(-.hat, -.sigma, -.fitted)
  my_list$observations <- model.metrics %>% count(class)
  my_list$shapiro <- model.metrics %>% shapiro_test(.resid)
  my_list$levene <- levene_test(.resid ~ class, data = model.metrics)
  my_list$outliers <- model.metrics %>% filter(abs(.std.resid) > 3)
  return(my_list)
}

# Run the assumptions function on the model
assumptions_list <- assumptions(adas_cog_lm)
assumptions_list
```
Note: NS.

### Education as covariate instead of MoCA
```{r}

adas_cog_lm_ed <- lm(adas_cog_plus_wins~class + sex + age + bmi + education + rct, final_data_lpa)
Anova(adas_cog_lm_ed, type = "III")
summary(adas_cog_lm_ed)
performance(adas_cog_lm_ed)
emmeans(adas_cog_lm_ed, ~ class)
pairs(emmeans(adas_cog_lm_ed, ~ class, adjust = "tukey"))

# Assumptions
## Function to summarize assumptions
assumptions <- function(x) {
  my_list <- list()
  model.metrics <- augment(x) %>% dplyr::select(-.hat, -.sigma, -.fitted)
  my_list$observations <- model.metrics %>% count(class)
  my_list$shapiro <- model.metrics %>% shapiro_test(.resid)
  my_list$levene <- levene_test(.resid ~ class, data = model.metrics)
  my_list$outliers <- model.metrics %>% filter(abs(.std.resid) > 3)
  return(my_list)
}

# Run the assumptions function on the model
assumptions_list <- assumptions(adas_cog_lm_ed)
assumptions_list
```
Note: NS.

## Raw
```{r}

adas_cog_lm <- lm(adas_cog_plus~class + sex + age + bmi + moca + rct, final_data_lpa)
Anova(adas_cog_lm, type = "III")
summary(adas_cog_lm)
performance(adas_cog_lm)
emmeans(adas_cog_lm, ~ class)
pairs(emmeans(adas_cog_lm, ~ class, adjust = "tukey"))

# Assumptions
## Function to summarize assumptions
assumptions <- function(x) {
  my_list <- list()
  model.metrics <- augment(x) %>% dplyr::select(-.hat, -.sigma, -.fitted)
  my_list$observations <- model.metrics %>% count(class)
  my_list$shapiro <- model.metrics %>% shapiro_test(.resid)
  my_list$levene <- levene_test(.resid ~ class, data = model.metrics)
  my_list$outliers <- model.metrics %>% filter(abs(.std.resid) > 3)
  return(my_list)
}

# Run the assumptions function on the model
assumptions_list <- assumptions(adas_cog_lm)
assumptions_list
```